Stage-structured evolutionary demography for simultaneous hermaphrodites with partial selfing

### 2020-09-03 notes/notes-SA-partSelf-wDemography.tex
updated notes with inv cond

{
\small
\begin{align*} \label{eq:AtildeCoexistence}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] &= \tilde{\mbf{U}} + \tilde{\mbf{F}}\\
		&= \left(
			\begin{array}{cc|cc|cc}
				\mcal{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mcal{U}^X_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\\hline
				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa} & \mbf{0} &  \mbf{0}\\\hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa}
			\end{array} \right) \\ 
		&+ \left(
			\begin{array}{cc|cc|cc}
				C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}  & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ 
				q^{\prime}_{A}  (1 - C_{AA}) \mbf{F}_{AA}& q^{\prime}_{A} (1 - C_{AA}) \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\
				q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA} & q^{\prime}_{a}(1 - C_{AA})  \mbf{F}_{AA} & \frac{1}{2}  (1 - C_{Aa}) \mbf{F}_{Aa}& \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & q^{\prime}_{A} (1 - C_{aa})\mbf{F}_{aa}  & q^{\prime}_{A}  (1 - C_{aa})\mbf{F}_{aa}\\ \hline
				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} &  C_{aa} (1 - \delta) \mbf{F}_{aa}\\ 
				\mbf{0} & \mbf{0} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa}) \mbf{F}_{Aa} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa})\mbf{F}_{Aa}& q^{\prime}_{a}(1 - C_{aa})  \mbf{F}_{aa} & q^{\prime}_{a} (1 - C_{aa})\mbf{F}_{aa} 
			\end{array} \right) \numberthis
\end{align*}
}

@@

{
\footnotesize
 \begin{align*} \label{eq:M22}
	\mbf{M}_{22} = 
		\frac{1}{\lambda_{AA}}\left(\begin{array}{ccc}
\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
 \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa} + \Theta\\
  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
			\end{array} \right), \numberthis
\end{align*}
}
where $\Theta = \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime$, and

@@

{
\footnotesize
\begin{align*} \label{eq:eigAA}
	\tilde{\zeta}_{AA} = 
		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{ccc}
\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
 \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa} + \Theta\\
  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
			\end{array} \right). \numberthis
\end{align*}
}

@@

where $\Phi = \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$.

@@

\newpage
\end{landscape}

>>>>

{
\small
\begin{align*} \label{eq:AtildeCoexistence}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] &= \tilde{\mbf{U}} + \tilde{\mbf{F}}\\
		&= \left(
			\begin{array}{cc|cc|cc}
				\mcal{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mcal{U}^X_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\\hline
				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa} & \mbf{0} &  \mbf{0}\\\hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^S_{aa} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
			\end{array} \right) \\ 
		&+ \left(
			\begin{array}{cc|cc|cc}
				C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}  & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ 
				q^{\prime}_{A}  (1 - C_{AA}) \mbf{F}_{AA}& q^{\prime}_{A} (1 - C_{AA}) \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\
				q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA} & q^{\prime}_{a}(1 - C_{AA})  \mbf{F}_{AA} & \frac{1}{2}  (1 - C_{Aa}) \mbf{F}_{Aa}& \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & q^{\prime}_{A} (1 - C_{aa})\mbf{F}_{aa}  & q^{\prime}_{A}  (1 - C_{aa})\mbf{F}_{aa}\\ \hline
				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} &  C_{aa} (1 - \delta) \mbf{F}_{aa}\\ 
				\mbf{0} & \mbf{0} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa}) \mbf{F}_{Aa} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa})\mbf{F}_{Aa}& q^{\prime}_{a}(1 - C_{aa})  \mbf{F}_{aa} & q^{\prime}_{a} (1 - C_{aa})\mbf{F}_{aa} 
			\end{array} \right) \numberthis
\end{align*}
}

@@

{
\footnotesize
 \begin{align*} \label{eq:M22}
	\mbf{M}_{22} = 
		\frac{1}{\lambda_{AA}}\left(\begin{array}{ccc}
\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
 \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa} + \Theta\\
  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
			\end{array} \right), \numberthis
\end{align*}
}
where $\Theta = \frac{(1 - C_{AA})}{p_{n}} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime$, and

@@

{
\footnotesize
\begin{align*} \label{eq:eigAAFull}
	\tilde{\zeta}_{AA} = 
		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{ccc}
\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
 \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa} + \Theta\\
  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
			\end{array} \right). \numberthis
\end{align*}
}

@@

where $\Phi = \frac{(1 - C_{aa})}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$.

\subsection*{Conditions for the maintenance of SA polymorphism}

My PhD supervisor Tim and I have a longstanding joke (originally referencing something Alexei Kondrashov used to say during lectures) for when, seemingly against all odds, your equations simplify to give you a beautiful result. We call it "a gift from Jesus", and here Jesus really is being generous! 
\bigskip

OK... let's start with a pretty general situation: we will solve for the conditions where SA polymorphism is maintained, defined by where $\tilde{\zeta}_{i} > 1$, where $i \in \{AA,aa\}$ (i.e., where the right hand side of Eq(\ref{eq:eigAAFull}) is greater than $1$). The expressions get pretty ugly, so for simplicity we will make two simplying assumptions. First, we will assume additive SA fitness effects ($h_f = h_m = 1/2$). Second, that at the boundary equilibria, the frequency of the $a$ allele in individuals produced by selfing is equal to the frequency in individuals produced by outrcrossing, i.e., $\hat{\mbf{p}}_{X,i} = \hat{\mbf{p}}_{S,i}$. For ease of comparison with previous population genetic models (Jordan \& Connallon 2014; Olito 2017), we solve $\tilde{\zeta}_{i} > 1$ for $s_f$. In this case, SA polymorphism will be maintained when

\begin{equation} \label{eq:InvCondSimp}
	\frac{s_m (1 - C)}{(1 - s_m) (1 + C (1 - 2 \delta)} < s_f < \frac{s_m (1 - C)}{1 + s_m (1 - C ) + C (1 - 2 \delta)}.
\end{equation}

\noindent Remarkably, these conditions are mathematically identical to those derived for additive SA fitness effects by Jordan \& Connallon (2014) and the 1-locus invasion condition derived in Olito (2017). Relaxing the above simplifying assumptions quickly leads to complicated expressions. However, we get very similar expressions to the pop.~gen.~models if we relax the assumption of additive fitness effects and allow for arbitrary dominance.

One scenario where we do get interesting behaviour is when we relax the assumption that $\hat{\mbf{p}}_{X,i} = \hat{\mbf{p}}_{S,i}$. Again, the expressions are complicated (not worth writing out here), but here is what happens: ($1$) when $\hat{\mbf{p}}_{X,i} > \hat{\mbf{p}}_{S,i}$, the funnel "opens up", so that polymorphism is maintained in a saddle-shaped area straddling the $1:1$ line such that the regions where either the $A$ or $a$ allele fixes shrink into the upper-left and lower-right hand corners of $s_f \times s_m$ parameter space. This dramatically increses the proportion of parameter space where polymorphism is maintained. ($2$) When $\hat{\mbf{p}}_{X,i} < \hat{\mbf{p}}_{S,i}$ the funnel shrinks such that polymorphism is only maintained in regions where $s_f,s_m > 0$, and gets progressively smaller until srong selection is required through both female and male function to maintain polymorphism.

Now, what's interesting (assuming I understand how $\hat{\mbf{p}}_{S,i}$ and $\hat{\mbf{p}}_{X,i}$ are calculated) is that $\hat{\mbf{p}}_{X,i} = \hat{\mbf{p}}_{S,i}$ unless both $C > 0$ and $\delta > 0$, in which case  $\hat{\mbf{p}}_{X,i} > \hat{\mbf{p}}_{S,i}$. So, I think this means that when there is selfing and some inbreeding depression, the parameter space where polymorphism is maintained is expanded relative to classic population genetic models. \hl{Need to think about this carefully. Interested to hear your thoughts Lotte!}

@@

\newpage
